\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{Assessing economic inequality with tax data - Switzerland from 1945 to 2010}

### 2014-05-25 paper/options/global_r_options.Rnw
fixes, so compiling Rnw works
merging minor diff between git and local

  # messages 
  opts_chunk$set(warning=FALSE)
  # code evaluation 
  opts_chunk$set(eval=TRUE)
  # code decoration
  opts_chunk$set(tidy=TRUE, prompt=TRUE, comment='##', highlight=TRUE, size='normalsize')
  # cache options
  opts_chunk$set(cache.path='cache/')  
  # Caching on by default
  opts_chunk$set(cache=T)
  # plot options 
  #if("tikzDevice" %in% rownames(installed.packages()) == TRUE) {
	# opts_chunk$set(fig.path='usr/graphics/dynamic/', dev='tikz', fig.keep='high', fig.show='hold')  
  #} else {
	# opts_chunk$set(fig.path='usr/graphics/dynamic/', dev='pdf', fig.keep='high', fig.show='hold')  
  #}

>>>>

  # RF: I turn source code off
  opts_chunk$set(echo=FALSE)

  # messages 
  opts_chunk$set(warning=FALSE)
  # code evaluation 
  opts_chunk$set(eval=TRUE)
  # code decoration
  opts_chunk$set(tidy=TRUE, prompt=TRUE, comment='##', highlight=TRUE, size='normalsize')
  # cache options
  opts_chunk$set(cache.path='cache/')  
  # Caching on by default
  opts_chunk$set(cache=FALSE)
  # plot options 
  #if("tikzDevice" %in% rownames(installed.packages()) == TRUE) {
	# opts_chunk$set(fig.path='usr/graphics/dynamic/', dev='tikz', fig.keep='high', fig.show='hold')  
  #} else {
	# opts_chunk$set(fig.path='usr/graphics/dynamic/', dev='pdf', fig.keep='high', fig.show='hold')  
  #}

### 2014-05-25 paper/paper.tex
merging minor diff between git and local

@@

@@

While there are several studies about income inequality in Switzerland, the publications concerning the distribution of wealth are scarcer, albeit the distribution of wealth seems to be a relevant dimension shaping the economic well-being of individuals as well. The cross-national Data center in Luxembourg is expanding their efforts, constructing the first cross-national wealth database. But, up to the day information for Switzerland is not available. 
% ab hier wird es irgnediwe unklar
Shorrock et al. (2013) unify several datasources (including tax data) to assess the pattern of global wealth. Their databook includes figures for Switzerland. In contrast to inequality of income concerning inequality of wealth Switzerland takes a leading position \\

% the credit suisse wealth report already uses wealth tax records, but they use it only for the period from 2000-2011. They did exactly the same for Denmark and Norway, which could be interesting to compare.
% Erbschaftssteuer erwähnen und bei den Global Wealth Reports der Credit Suisse und den Studien im Forschungsantrag weitermachen.

Up to the day, Switzerland can be situated according to the actual level of income inequality in western societies as there is a huge effort to collect data which can be harmonized to comparable measures (see Luxembourg income study). However, it is unclear how the bias through non-response affects the overall measure of inequality. Likewise a long and consistent time-series allowing to identify and explore development patterns on every point of the distribution (not only the top-shares) and a time series for inequality of wealth is missing. Building on recent developments in the field of inequality research, we assess the suitability of the publicly accessible tax data to report inequality and its changes over time. First of all this includes a discussion of the accessible measures in context of a reflection about the state of the art conceptualization of economic resources as an indicator for economic well-being. Second, we summarize and apply tax data specific technics to construct suitable measures of inequality. We expand the given set by an in depth discussion of a newly applied step to handle the incomplete coverage in tax data statistics. Third, we integrate the development of Swiss inequality into the international picture.

@@

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{verbatim}
## Wald test for kanton:zeros
##  in lm(formula = G_steink ~ kanton + kanton:zeros + factor(steuerperiode) + 
##     factor(steuerperiode):zeros, data = df)
## F =  3.282  on  27  and  262  df: p= 4e-07
\end{verbatim}
\end{kframe}
\end{knitrout}

>>>>

% abstract müsste angepasst werden.
% Wir machen nichts mehr zu Vermögen (obwohl der Trend auch dahin geht, dass neben dem Einkommen auch das Vermögen eine zentrale Untersuchunsgrösse des inviduellen Wohlstandss darstellt)

% Alternative
% There is empirical evidence that economic inequality increased in the majority of western countries over the last decades (OECD, 2011; Gornick and Jäntti, 2013). In Switzerland, however, the development is unclear, as there is evidence for trends in both directions. Part of the inconclusive picture is due to different methodological approaches. In this paper we discuss the role of taxdata concerning the assessment of inequality in income. The focus of the discussion lays herein to show the benefits and shortcomings of tax data compared to current state of the art measurement concepts of economic inequality. We present common and new strategies to handle tax data specific methodological difficulties and compare results out of aggregated federal tax statistics to results from the Household Budget Survey (HBS). We can show to which extend survey data underestimate inequality in income. Following the results out of the taxdata Switzerland experienced in slight rise in inequality in recent years, similar to other western countries, but only because of rise in upper percentiles of the income distribution.

% Salverda-Stuff und das neue Piketty Buch sollten irgendwo erwähnt werden
%

@@

% Bei SILC einsteigen. Allenfalls eine Grafik bringen mit den unterschiedlichen Reihen.
% Daraus wird auch ersichtlich, dass keine der verfügbaren Reihen wirklich länge Zeiträume abdecken
% we should mention that Eurostat uses EU-SILC alike and show, that there is a time series for SILC-Data aswell. Additionally we should mention that the Federal Statistical Office uses the HABE statistic

% Allenfalls erwähnen, dass die Schweiz nicht Teil des GINI database ist.
http://www.gini-research.org/articles/data_2

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=\maxwidth]{figure/Swiss_inequality_Plot}

@@

While there are several studies about income inequality in Switzerland, the publications concerning the distribution of wealth are scarcer, albeit the distribution of wealth seems to be a relevant dimension shaping the economic well-being of individuals as well. The cross-national Data center in Luxembourg is expanding their efforts, constructing the first cross-national wealth database. But, up to the day information for Switzerland is not available.

% ab hier wird es irgnediwe unklar
Shorrock et al. (2013) unify several datasources (including tax data) to assess the pattern of global wealth. Their databook includes figures for Switzerland. In contrast to inequality of income concerning inequality of wealth Switzerland takes a leading position \\

% Albeit the awarness of an assessment of income and wealth simuntanously is rising, we focus in this paper on income inequality, which is the know as the core driver of economic inequlity
% Dann lassen sich einige Passagen streichen. Man könnte auch auf verschiedene Publikationen verweisen, welche die Bedeutung von Vermögen untersuchen -> Buch von Piketty, Müller und Schoch und dann aber ziemlich schnell darauf verweisen, dass wir uns hier auf Einkommen konzentrieren.
% Auch erste Dell Studie erwähnen -> es hat Zahlen zur Schweiz. Auch Probleme erwähnen (Fehlende Informationen zu Altersvorsorgevermögen -> hat aber niemand eine Lösung bisher.)
% Studies on inequality have to adress several thorny challenges. One has to define the centrale measure of inequality and find adquate data. Recent studies accentuate the importance of a broader concept of economic wellbeeing including measures for income, consumption and wealth alike. But each area is confronted with own measurement problems. Measure of consumptions are largely missing in tax data, albeit deductions can be understood as mandatory consumptions somehow. Measures of wealth are included in tax data, but are not connectable to the respctive income and therefore can only be analyzed separate. In danger of falling a pittraps it seems appriate to move in small steps. Therefore we focus our analysis on income, which is undoubtly a crucial indicator of economic wellbeeing.
% Also the Federal Statistical Office is increasing their effort to get sound data concering the distribution of wealth vgl. Müller und Schoch (2014).

% Zusammenfassung Stand der Forschung und festlegen unserer Ausrichtung

Up to the day, Switzerland can be situated according to the actual level of income inequality in western societies as there is a huge effort to collect data which can be harmonized to comparable measures (see Luxembourg income study, EU-SILC). However, it is unclear how the bias through non-response affects the overall measure of inequality. Likewise a long and consistent time-series allowing to identify and explore development patterns on every point of the distribution (not only the top-shares) and a time series for inequality of wealth is missing. Building on recent developments in the field of inequality research, we assess the suitability of the publicly accessible tax data to report inequality and its changes over time. First of all this includes a discussion of the accessible measures in context of a reflection about the state of the art conceptualization of economic resources as an indicator for economic well-being. Second, we summarize and apply tax data specific technics to construct suitable measures of inequality. We expand the given set by an in depth discussion of a newly applied step to handle the incomplete coverage in tax data statistics. Third, we integrate the development of Swiss inequality into the international picture.

% delete passages about wealth and include passage about the comparison with survey data

@@

% Section from an older version. Currently disabled.

@@

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# \\\#\# Attaching package: 'survey'\\\#\# \\\#\# Das folgende Objekt ist maskiert from 'package:graphics':\\\#\# \\\#\#\ \ \ \  dotchart}}\begin{verbatim}
## Wald test for kanton:zeros
##  in lm(formula = G_steink ~ kanton + kanton:zeros + J + zeros:J, 
##     data = df)
## F =  3.282  on  27  and  262  df: p= 4e-07
\end{verbatim}
\end{kframe}
\end{knitrout}
